---
title: "Multinomial Logistic Regression on Speed Dating Data"
author: "Roko Mijic"
date: "23 May 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(zoo)
library(dplyr)
library(readr)
library(corrplot)
library(ggplot2)
library(glmnet)

```

Read the data

```{r}
setwd('C:/Users/rmiji/OneDrive/R_working_dir/Signal')
df = read.csv('./Week4/speeddating-aggregated.csv')
```

Find most common careers and select predictors, reorder dataframe:

```{r}
common_careers = as.numeric(names(sort(table(df$career_c),decreasing = TRUE)[1:4]))
df = (dplyr::select(df, -gender, -race, -race_o, -dec_o))[,c(1:5,7:23,6)]
```

filter 


```{r}

df = dplyr::filter(df, career_c %in% common_careers)

predictors = scale(data.matrix(df[,1:ncol(df)-1]))

target = data.matrix(df[,ncol(df)])

```

create a multinomial model which predicts occupation, then visualise the coefficients of that model

```{r}

multinom_model = glmnet(predictors, target, family="multinomial", lambda = 0)

coefmat = as.matrix(do.call(cbind, coef(multinom_model) ))[-1,]

colnames(coefmat) = names(coef(multinom_model))
colnames(coefmat) = c("Lawyer", "Academic", "Arty-farty", "business/finance")

corrplot(coefmat, is.corr = FALSE)

```


Now make predictions using the model, and find the principal components of the predicted categories.

```{r}

predicted_log_odds = (predict(multinom_model, predictors, s = 0  ))[,,1]

prc_obj = prcomp(predicted_log_odds)

change_of_basis = prc_obj$rotation

prc_obj$sdev

row.names(change_of_basis) = c("Lawyer", "Academic", "Arty-farty", "business/finance")

```

It turns out that the biggest principal component seems to correspond to people who have highly paid jobs versus those who don't, and the second component corresponds to mathiness:

```{r}

colnames(change_of_basis) = c("Poorly paid", "Less math oriented", "???", "noise")

corrplot(change_of_basis)

```


We can write a function that converts a row of this matrix of predictions (in this case corresponding to a particular person who went speed dating), and turns it from log odds into probabilities


```{r}
preds = matrix(data = c(1,2,3,4), nrow = 2)
preds
rownum = 1

probabilities = function(preds, rownum)
{
  proportional_p_row = exp(preds[rownum,])
  normalizer = sum(proportional_p_row) 
  return(proportional_p_row/normalizer)
}

probabilities(predicted_log_odds,1)

```


