---
title: "Factor_Analy"
author: "Roko Mijic"
date: "17 May 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(corrplot)
library(dplyr)
library(psych)
set.seed(1) 
X = rnorm(100)

mean(X)

Y = rnorm(100)
Z = rnorm(100)

factors = data.frame(X,Y,Z)

```


A function can be written which creates noisy indicators for vectors:


```{r}
noisyIndicators = function(feature, k, correlation)
{
  ls = vector(mode = "list", length = k)
  for(i in c(1:k))
  {
    indicator = feature - rnorm(length(feature),mean = 0, sd = (1-correlation^2)^0.5  )
    ls[[i]] = indicator
  }
  
  df = data.frame(ls)
  colnames(df) = c(1:k)
  return(df)  
}

testvec = (c(1,2,3,4,5) - rep(mean(c(1,2,3,4,5))))/sd(c(1,2,3,4,5))
testvec
noisyIndicators( testvec,5,0.1)

```

Now we will create noisy indicators for our normal random variables X and Y

```{r}
p=0.9
noisies = data.frame(noisyIndicators( X,4,p), noisyIndicators( Y,3,p))
cor(noisies)
corrplot(cor(noisies))

```

now run PCA on the noisies and factors. They are what we expect:

```{r}

pr_noise = prcomp(noisies)
corrplot(cor(pr_noise$x, noisies))
corrplot(cor(pr_noise$x, factors))

```

How does this differ from orthogonal factor analysis?

```{r}

fa_noise = fa(noisies, nfactors = 2, rotate = "varimax")
corrplot(cor(fa_noise$scores, factors))

```

Quite a bit closer, right?

Now we will generate 50 variables..

```{r}
vars = vector(mode = "list", length=50)
for (i in seq(50)) {
  vars[[i]] = X*runif(1) + Y*runif(2) + Z*runif(3) + 0.5*rnorm(1)
}
vars = data.frame(vars)
colnames(vars) = seq(50)
```

..and run PCA and factor analysis on them. Additionally, we will compute the correlations between (first 3 principal components for PCA) / 3 modeled factors and X,Y,Z.

```{r}

pr_vars = prcomp(vars)
corrplot(cor(pr_vars$x[,1:3], factors))

fa_vars = fa(vars, nfactors = 3, rotate = "varimax")
corrplot(cor(fa_vars$scores, factors))

```

Let's play with oblique factor analysis. First, set W and check the correlation between W and Y.

```{r}
W = 0.5*X + Y
cor(W, Y) # 0.905

```

Then generate 10 noisy indicators and plot the associated correlation matrix.

```{r}
p = 0.8
noisies = data.frame(noisyIndicators(X, 10, p), noisyIndicators(W, 4, p))

corrplot(cor(noisies))


```

Let's now compare orthogonal vs oblique factor analysis:

```{r}
ortho_fa = fa(noisies, nfactors = 2, rotate = "varimax")
oblq_fa = fa(noisies, nfactors = 2, rotate = "oblimin")

corrplot(cor(ortho_fa$scores, noisies))
corrplot(cor(ortho_fa$scores, factors))

corrplot(cor(oblq_fa$scores, noisies))
corrplot(cor(oblq_fa$scores, factors))
```