---
title: "OKCupid_data"
author: "Roko Mijic"
date: "18 May 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(psych)
library(zoo)
library(dplyr)
library(readr)
library(corrplot)
library(ggplot2)

setwd('C:/Users/rmiji/OneDrive/R_working_dir/Signal')
source('./Week2/resampling_utils.R')
df = readRDS('./Week3/parsed_data.rds')
qdata = read_delim('./Week3/question_data.csv', ';')

```

Here we will process the OKCupid data. First read the OKCupid data. The data has 68371 rows and 2624 columns, about 179,000,000 values. 76% of the values are "NA". 
We can reduce the data by taking a random sample of 10% of the rows where gender = male. This can be repeated for females. 

```{r message=FALSE}

df = dplyr::filter(df, d_gender == "Man" )

df = reduce_data(df, frac = 0.80)

Total_vals = dim(df)[1]*dim(df)[2]
Total_vals
frac_na = sum(is.na(df))/Total_vals
frac_na
```

Do some minor renaming of the question data:

```{r message=FALSE}
colnames(qdata)[[1]] = "q_num"
rownames(qdata) = paste("q", rownames(qdata), sep = "" )

```

Now we will process the question data to find question numbers with two answers, three answers and four answers

```{r message=FALSE}
q_ans_2 = dplyr::filter(qdata, is.na(option_3), is.na(option_4))$q_num
q_ans_3 = dplyr::filter(qdata, !is.na(option_3), is.na(option_4))$q_num
q_ans_4 = dplyr::filter(qdata, !is.na(option_3), !is.na(option_4))$q_num
```


We write a function which automatically turns a dataframe of binary factor features into numerical ones, another function to remove constant columns, and a final one to turn question numbers into descriptive question texts: 

```{r message=FALSE}
# function to make a non-numeric dataframe into a numeric one
numericalize = function(x){  dm = data.matrix(x) ;  class(dm) = "numeric" ;  return(data.frame(dm-1)) }

# function to remove constant columns
deconstantize = function(x) {x[,apply(x, 2, var, na.rm=TRUE) != 0]}

# function to turn a set of question numbers into a set of descriptive question texts 
descriptiveize = function(x) {return(qdata[match(x, qdata$q_num),2]) }

```

Now we get dataframes that are the data with just certain types of question included, for example two answer questions, three answer questions, etc: 

```{r message=FALSE}
df_qs = dplyr::select( df, starts_with("q"))
df_ds = dplyr::select( df, starts_with("d"))
df_ps = dplyr::select( df, starts_with("p"))
df_2ansq = df_qs[, names(df_qs) %in% q_ans_2]
df_3ansq = df_qs[, names(df_qs) %in% q_ans_3]
df_4ansq = df_qs[, names(df_qs) %in% q_ans_4]
df_misc = dplyr::select( df, -starts_with("q"), -starts_with("p"), -starts_with("d") )
```

Turn the two-answer questions into 0-1 numericals:

```{r message=FALSE}
df_2ansq = numericalize(df_2ansq)
```

We can remove costant columns:

```{r message=FALSE}
df_2ansq = deconstantize(df_2ansq)
df_2ansq[1:5, 1:5]
```

compute the correlation matrix using "pairwise.complete.obs", then find eigenvalues and eigenvectors

```{r message=FALSE}
corM = cor(df_2ansq, use = "pairwise.complete.obs"  )
corM[is.na(corM)] = 0
```

(*)

```{r message=FALSE}
eigenobj = eigen(corM, symmetric = TRUE)
rownames(eigenobj$vectors) = rownames(corM)

```

Examine the large eigenvalues, make a scree plot. Arbitrarily choose eigenvalue magnitude > 11.85 (this needs tinkering with using the scree plot)

```{r message=FALSE}

large_evals_mask = abs(eigenobj$values) > 11.85

large_evals_abs = sort(abs((eigenobj$values)[large_evals_mask]), decreasing = TRUE)

large_evecs = (eigenobj$vectors)[,large_evals_mask]
eval_df = data.frame(eigenvector_number = c(1:length(large_evals_abs)), eigenvalue = large_evals_abs)

ggplot() + geom_point(data = eval_df  , aes(x = eigenvector_number, y=eigenvalue ) )


corM[1:5,1:5]

print_evec = function(n){}
  
print(descriptiveize(names(sort(large_evecs[,1]))[1:10]))
print(descriptiveize(names(sort(large_evecs[,2])[1:10])))
print(descriptiveize(names(sort(large_evecs[,3])[1:10])))
print(descriptiveize(names(sort(large_evecs[,4])[1:10])))
print(descriptiveize(names(sort(large_evecs[,5])[1:10])))
print(descriptiveize(names(sort(large_evecs[,6])[1:10])))
print(descriptiveize(names(sort(large_evecs[,7])[1:10])))

```



We can also run factor analysis from the line (*) :

```{r message=FALSE}
fact_obj = fa(corM, nfactors = 17, rotate="oblimin")

fact_obj$values



head(fact_obj$loadings)

(fact_obj$loadings)[1:10,1:10]

fact_obj$fit



```












