---
title: "OKCupid_data"
author: "Roko Mijic"
date: "18 May 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

source('./Week2/resampling_utils.R')
library(zoo)
library(dplyr)
library(readr)
library(corrplot)
library(ggplot2)

```

read the OKCupid data. The data has 68371 rows and 2624 columns, about 179,000,000 values. 76% of the values are "NA". 
We can reduce the data by taking a random sample of 3% of the rows. 

```{r}

df = readRDS('./Week3/parsed_data.rds')
qdata = read_delim('./Week3/question_data.csv', ';')
rownames(qdata) = paste("q", rownames(qdata), sep = "" )
rownames(qdata) 

df = reduce_data(df, frac = 0.2)
dim(df)

Total_vals = dim(df)[1]*dim(df)[2]
Total_vals

frac_na = sum(is.na(df))/Total_vals
frac_na
```

Now we will process the question data to find question numbers with two answers, three answers and four answers

```{r}
colnames(qdata)[[1]] = "q_num"
head(qdata)

q_ans_2 = dplyr::filter(qdata, is.na(option_3), is.na(option_4))$q_num
q_ans_3 = dplyr::filter(qdata, !is.na(option_3), is.na(option_4))$q_num
q_ans_4 = dplyr::filter(qdata, !is.na(option_3), !is.na(option_4))$q_num
```



We write a function which automatically turns a dataframe of binary factor features into numerical ones, and another function to remove 

```{r}

numericalize = function(x){  dm = data.matrix(x) ;  class(dm) = "numeric" ;  return(data.frame(dm-1)) }
deconstantize = function(x) {x[,apply(x, 2, var, na.rm=TRUE) != 0]}
descriptiveize = function(x) {return(qdata[rownames(qdata) %in% x,2]) }

```

Now we get dataframes that are the data with just certain types of question included, for example two answer questions, three answer questions, etc: 

```{r}

df_qs = dplyr::select( df, starts_with("q"))
df_ds = dplyr::select( df, starts_with("d"))
df_ps = dplyr::select( df, starts_with("p"))
df_2ansq = df_qs[, names(df_qs) %in% q_ans_2]
df_3ansq = df_qs[, names(df_qs) %in% q_ans_3]
df_4ansq = df_qs[, names(df_qs) %in% q_ans_4]
df_misc = dplyr::select( df, -starts_with("q"), -starts_with("p"), -starts_with("d") )
users = dplyr::select( df, d_username ) 
```

Turn the two-answer questions into 0-1 numericals:

```{r}
df_2ansq = numericalize(df_2ansq)
```

We can remove costant columns:

```{r}
df_2ansq = deconstantize(df_2ansq)

df_2ansq[1:15, 1:15]
```


We can impute missing values simplistically using na.aggregate from zoo

```{r}
#df_2ansq = na.aggregate(df_2ansq)
df_2ansq[1:15,1:15]
```

compute the correlation matrix using "pairwise.complete.obs", then find eigenvalues and eigenvectors

```{r}

head(df_2ansq)

corM = cor(df_2ansq, use = "pairwise.complete.obs"  )
corM[is.na(corM)] = 0
eigenobj = eigen(corM, symmetric = TRUE)
rownames(eigenobj$vectors) = rownames(corM)

```

Examine the large eigenvalues, make a scree plot. Arbitrarily choose eigenvalue magnitude > 12 (this needs tinkering with using the scree plot)

```{r}

large_evals_mask = abs(eigenobj$values) > 12
large_evals = abs((eigenobj$values)[large_evals_mask])
large_evecs = (eigenobj$vectors)[,large_evals_mask]
eval_df = data.frame(eigenvector_number = c(1:length(large_evals)), eigenvalue = large_evals)

ggplot() + geom_point(data = eval_df  , aes(x = eigenvector_number, y=eigenvalue ) )


corM[1:15,1:15]

names(prcomp_obj$sdev) = c(1:length(prcomp_obj$sdev))
compdata = data.frame(compnum = c(1:length(prcomp_obj$sdev)), sdev = prcomp_obj$sdev)
ggplot() + geom_point(data = compdata[1:20,] , aes(x = compnum, y=sdev) )


print(descriptiveize(names(sort(large_evecs[,1])[1:50])))
print(descriptiveize(names(sort(large_evecs[,2])[1:50])))
print(descriptiveize(names(sort(large_evecs[,3])[1:50])))


```










