---
title: "PCA on a speed dating dataset"
author: "Roko Mijic"
date: "16 May 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load required packages, load the data, remove NA values:

```{r echo=TRUE, message=FALSE}
library(dplyr)
library(zoo)
library(psych)
library(ggplot2)
library(corrplot)
library(DAAG)
df = msq
df = dplyr::select(df, active:scornful)
df = na.aggregate(df)
```

Now get a principal component object for the data

```{r message=FALSE}
prc = prcomp(df, scale=TRUE)
compsds = data.frame(comp = c(1:75), sdevs = prc$sdev)
```


write a function to output the top 5 components of the nth principal component


```{r message=FALSE}
top = function(n) 
{
  nthcomp = prc$rotation[,n]
  sorted_nth = nthcomp[order(abs(nthcomp), decreasing = TRUE)] 
  return(sorted_nth[1:5])
}
top(1)
top(2)
top(3)
top(4)
top(5)
```

plot the standard deviations of the components

``` {r message=FALSE}
ggplot(compsds, aes(x = comp, y = sdevs)) + geom_point( )

```


use the first n principal components to predict extraversion and neuroticism. We'll reload the data first. 

```{r message=FALSE}
df = msq
df_targets = na.aggregate(dplyr::select(df,  Extraversion, Neuroticism))
df_preds = na.aggregate(dplyr::select(df, active:scornful))
princo = prcomp(df_preds, scale=TRUE)
numfeats = ncol(df_preds)
n=10
df_projected_extra = dplyr::select(data.frame((princo$x)[,1:n], df_targets), -Neuroticism)
df_projected_neuro = dplyr::select(data.frame((princo$x)[,1:n], df_targets), -Extraversion)

head(df_projected_extra)
```

Plot extraversion and neuroticism as a function of the first two principal components

```{r message=FALSE}
ggplot() + geom_point(data=df_projected_extra, aes(x=PC1, y=PC2, colour = Extraversion)) + scale_colour_gradientn(colours=rainbow(4))
```

get the goodness of a fit vs the number of components used

```{r message=FALSE, warning=FALSE}
sumss = 0
fit_n_comps = function(n)
{
  
  df_projected_extra = dplyr::select(data.frame((princo$x)[,1:n], df_targets), -Neuroticism)
  df_projected_neuro = dplyr::select(data.frame((princo$x)[,1:n], df_targets), -Extraversion)
  head(df_projected_extra)
  
  sumss = 0
  cvlm_extra = CVlm(data=df_projected_extra, Extraversion ~ ., plotit=FALSE,  printit = FALSE)
  R2_extra = cor(cvlm_extra$Extraversion, cvlm_extra$cvpred  )

  
  sumss = 0
  cvlm_neuro = CVlm(data=df_projected_neuro, Neuroticism ~ ., plotit=FALSE,  printit = FALSE)
  R2_neuro = cor(cvlm_neuro$Neuroticism, cvlm_neuro$cvpred  )
  
  
  return(list("R2_extra" = R2_extra,"R2_neuro" =  R2_neuro))

}

extra_R2s = c(1:numfeats)
neuro_R2s = c(1:numfeats)

for(n in c(1:numfeats))
{
  
  fits_obj = fit_n_comps(n)
  extra_R2s[[n]] = fits_obj[["R2_extra"]]
  neuro_R2s[[n]] = fits_obj[["R2_neuro"]]  
}


R2s = data.frame( "number_of_components" = c(1:numfeats), extra_R2s, neuro_R2s )

head(R2s)

ggplot() + geom_point(data =R2s, aes(x=number_of_components, y=extra_R2s), color = "Red") + 
           geom_point(data =R2s, aes(x=number_of_components, y=neuro_R2s), color = "Purple")


```






